---

---
title: "Homework 3 - ENVS 193DS"
author: "Siaa Singh"
format: html
editor: visual

[View this project on GitHub](https://github.com/siaasingh/ENVS-193DS_homework-03)
---

```{r}
library(tidyverse)
library(gt)     
library(janitor)  
library(readxl)



```

```{r}
install.packages("here")
library(here)
```

```{r}
# Read in the personal dataset
personal_data <- read_csv(here("..", "data", "ENVSDS_personaldatacsv.csv"))


```

---
Problem 1 
a. I could calculate the average number of hours I spent outdoors for each 
day of the week and compare those means. This would help me understand 
how my time outdoors varies by weekday—perhaps I spend more time outside on 
weekends because I have fewer classes, or less on weekdays with heavy homework 
loads like Monday or Wednesday.

---

b\. Visualization

```{r}
# Clean column names to simplify plotting
personal_data <- personal_data |> janitor::clean_names()

# Summarizing mean outdoor hours by weekday
summary_data <- personal_data |> 
  group_by(days_of_the_week) |> 
  summarize(avg_outdoors = mean(hours_spent_outdoors))

# Plot both individual points + summarized means
# Create a plot showing both raw data points and summarized averages:
# - geom_jitter() plots each individual observation, slightly offset to avoid overlap
# - geom_col() overlays bars showing the average hours spent outdoors for each weekday
ggplot(personal_data, aes(x = days_of_the_week, y = hours_spent_outdoors)) +
  geom_jitter(width = 0.2, color = "#1f78b4", alpha = 0.7) +  # raw data points and 
  #changing colours
  geom_col(data = summary_data, aes(y = avg_outdoors), 
           fill = "#33a02c", alpha = 0.6, width = 0.6) + #average (mean) values per weekday
  labs(
    title = "Average Hours Spent Outdoors by Day of the Week",
    x = "Day of the Week",
    y = "Hours Spent Outdoors"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", size = 14)
  )
```

c\. This figure shows the average number of hours spent outdoors by day of the week, with individual data points shown for each entry. The plot highlights that I tend to spend more time outdoors on weekends like Friday and Saturday, compared to weekdays such as Monday or Thursday. This may reflect differences in workload and class schedules.
